Online (Bandit) Policy Selection for EASY-Backfilling

### 2017-03-30 paper/paper.tex
incorporate nightly figures.
denis changes

Plan :
\begin{itemize}
\item Context - message : les machines sont de + en + complexes, les systemes
de management restent rudimentaires face a la complexite croissante de
l'architecture et des applications.
\item les habitudes et usages existants : justification du FCFS-EASY-BF,
developper rapidement ce qui existe
\item ce que l'on propose
\end{itemize}

Providing the computing infrastuctures needed to solve actual complex problems
arising in the various fields of the modern society (including climate change,
health, green energy or security) is a strategic challenge.
In particular, High Performance Computing (HPC) systems are evolving to
extreme-scale parallel and distributed platforms.
The race for always more computing power and storage capacity does not only
lead to sophisticated specific exascale platforms, but the objective of the
community is to design efficient sustained Petascale platforms.
However, there is still a long way to go and many scientific and technical
problems to solve by the system management software in order to adapt to such
large-scale evolution and the increasing complexity
in both architecture and applications.
The existing job and resource management sofwares allow to run tens thousands
jobs on hundreds thousands cores.
They are based on robust and rather simple policies.

\bigskip
The management generates a huge amount of data.
During the last few years, there was an explosion of the number of of works at
the interface of HPC and bigdata, dealing with learning algorithms
(including the authors of this paper~\cite{learningruntimes}).
Most of these studies target the process to determine more or less accurately
the value of some specific key parameters, with the idea of
learning better estimates should improve the performances of the resource
manager.

\bigskip
Our idea within this work is to learn how the management system behave within
the classical framework of of selecting jobs of the submission queue under
EASY-backfilling.
Contrarily to most existing studies which consider only one execution trace, we
extend the learning on a multiple trace basis.
Indeed, learning for a specific trace leads to limited results, customized for
one trace on a given platform (what happens if we target another platform?).
Moreover, this is usually inefficient to deal with extreme unpredictable
events.

@@

@@

This section presents related works in the area of scheduling for batch scheduling.

\subsection{Scheduling heuristics in HPC platforms}

While parallel job scheduling is a well studied theoretical
problem~\cite{leung2004handbook}, the practical ramifications, varying
hypotheses, and inherent uncertainty of the problem in HPC have driven
practitioners and researchers alike to use and study simple heuristics.
The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} Backfilling.

While Conservative Backfilling offers many advantages~\cite{bfchar}, it has a
significant computational overhead, perhaps explaining why most of the machines
of the top500 ranking~\cite{top500} still use at the time of this publication a
variant of EASY Backfilling.

\subsection{EASY}
There is a large body of work seeking to improve EASY. Indeed, while the
heuristic is used by various resource and job management softwares (most notably
SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by system
administrators.

Several works explore how to tune EASY by reordering waiting
and/or backfilling queues~\cite{Tsafrir_easypp_2005}, sometimes even in a
randomized manner~\cite{1592720}, as well as some
implementations~\cite{Jackson2001}. However, as successful as they may be, these works
do not address the dependency~\cite{variability} of scheduling metrics on the
workload. Indeed these studies most often report \textit{post-hoc} performance since
they compare algorithms after the workload is known.

The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
method to tuning these queues, although it requires simulated scheduling runs
at decision time and therefore costs much more than the natural execution of EASY.

\subsection{Data-aware resource management}

There is a recent focus on leveraging the high amount of data available in
large scale computing systems in order to improve their behavior. Some works
use collaborative filtering to colocate tasks in clouds by estimating
application interference~\cite{7516031}.  Others are closer to the application
level and use binary classification to distinguish benign memory faults from
application errors in order to execute recovery algorithms (see
~\cite{fmodeling} for instance).

Several works take a similar approach in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.

However, these works do not address the duality between the cumulative and
maximal scheduling costs, as mentionned in~\cite{learningruntimes}.

A recent work~\cite{mlmem} proposes to help
users specify better memory requirements by applying supervised learning
techniques to job metadata.

While these previous works intend to estimate uncertain parameters, they lack
in validating the impact on the system. We consider in this paper a more
pragmatic approach, which is to directly learn a scheduling policy from a given
seach space.

\subsection{Multi-Armed Bandits}

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards. At every round, an action (an arm) must be chosen
in a fixed set and the corresponding reward is observed. The goal of a MAB
algorithm is to maximize the total reward obtained in a number of rounds.
There are a number of works that address this problem under a variety of
constraints, The two most popular settings are the original stochastic  case,
and the adversarial case. See ~\cite{thompson} for the original work on the
stochastic case and ~\cite{Auer2002} for the UCB family of algorithms. The work
of ~\cite{Banos} is the earliest known work to us for the adversarial case
see~\cite{nonstoch} for the EXP family of algorithms. We refer to the
review~\cite{bubnow} for a comprehensive overview of the field. While these
algorithms bound the cumulative difference in loss to the best arm (the
regret), they have functional constraits such as the fact that the rewards
should be contained in a range. The\cite{Auer2002} heuristic called
Epsilon-Greedy is known to achieve good results in practice, and does not have
this requirement.

@@

%wait mwait lpf spf sqf lqf expfact mexpfact lrf srf laf saf
\begin{enumerate}
  \item FCFS: First-Come First-Serve, which is the widely used default policy~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  %\item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
  \item LPF: Longest estimated Processing time First.
  \item LQF: Largest resource requirement $q_j$ First.
  %\item SQF: Smallest resource requirement First.
  \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
    factor is defined as follows:
  \begin{equation} \frac{start_j - r_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
  where $start_j$ is the starting time of job $j$.
  \item SEXP: Smallest Expansion Factor First
  \item LRF: Largest Ratio $\frac{p_j}{q_j}$ First
  \item SRF: Smallest Ratio First
  \item LAF: Largest Area $ p_j \times q_j$ First
  %\item SAF: Smallest Area First
\end{enumerate}

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[scale=0.6]{variability.pdf}
  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (pre-processing described in subsection~\ref{sub:traces}) for a few policies.}
  \label{fig:mosn}
\end{figure*}

@@

Readers interested in replicating the experiments (or a part thereof) are
invited to peruse to the 'artifacts description' appendix of this paper.  We
provide to persistent archives containing the simulator used, the analysis code
and the associated workflow necessary to replicate all experiments and figures
present in this paper. All these parts's dependencies are automatically managed
via Nix~\cite{nix} and we invite the reader to replicate the results.

@@

\begin{figure*}[h]
  \centering
  \includegraphics[scale=0.6]{full-CTC-SP2.pdf}
  \includegraphics[scale=0.6]{full-KTH-SP2.pdf}\\
  \includegraphics[scale=0.6]{full-CEA-Curi.pdf}
  \includegraphics[scale=0.6]{full-SDSC-BLU.pdf}\\
  \includegraphics[scale=0.6]{full-SDSC-SP2.pdf}
  \includegraphics[scale=0.6]{full-ANL-Intr.pdf}

@@

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{UniLu-Ga.pdf}

  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$. The
    average is obtained by resampling the original trace 100 times. The dashed
    lines represent the 10th and 90th percentiles of the values across this
  resampling.}

@@

\clearpage
\section*{Artifacts Description}

In this annex, we describe the computational artifacts associated with this
work. We take a lightweight approach to replicability by packaging a workflow
that generates all the figures from this paper and making it available
publicly. Moreover, questions regarding replication can be freely directed at
the authors via e-mail.

@@

There are various approaches for making computational experiments replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach using the Nix~\cite{nix} packaging system.  All the
dependencies of our experiments (including our own code, dependencies for data
processing and visualization and workflow engine) are automatically managed by
Nix up to the actual execution of the code.

Running the experiments can be done on any platform equipped with the nix
workflow system by running:

\begin{lstlisting}
nix-build (fetchTarball ...) -A banditSelection
\end{lstlisting}

@@

 The experiments in this paper can run via in less than 24 hours on a large
 machine. More precisely, we had access to a Dell PowerEdge R730 equipped with
 a total of 56 threads @2.4GHz each and 757G of RAM.  For this reason, the
 experimental campaign was designed to be executed on a single host.

@@

 All experiments are tied together using Zymake~\cite{zymake}, a minimalistic
 workflow system designed for computational experiments. This system is
 analogous to a traditional build system with added workflow capabilities. The
 entire workflow that generates this article from the input data is contained
 in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
 found at the root of the main archive. This workfloww is composed of the
 following steps:

 \begin{itemize}

   \item[Preprocessing]: Data extraction from archives, filtering, resampling.
     This is principally using shell scripts for data manipulation (see file
     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter| for the
     filtering steps used) and ocaml code that implements the resampling method
     described in~\ref{sub:resampling}.

@@

   \item[Analysis]: This step runs mostly R code that generates the figures
     presented in this paper.
 \end{itemize}
\end{document}

>>>>

\subsection{Context}

Providing the computing infrastuctures needed to solve actual complex 
problems arising in the various fields of the modern society (including 
climate change, health, green energy or security) is a strategic challenge.
The main pillar to address this challenge is to build extreme-scale 
parallel and distributed platforms.
The race for always more computing power and storage capacity does not 
only lead to sophisticated specific exascale platforms, but the 
objective of the community is to design efficient sustained Petascale 
platforms.
However, there is still a long way to go and many scientific and 
technical problems to solve by the system management software in order 
to adapt to such large-scale evolution and the increasing complexity
in both architecture and applications.
The existing job and resource management softwares allow to run tens 
thousands jobs on hundreds thousands cores.
They are based on robust policies, however, despite their positive 
properties of relative simplicity and their guarantee of no job starvation,
there is still place for improvements.
We propose here to study how to tune the classical EASY-BackFilling 
heuristic using two new learning approaches (namely, .

\bigskip
The management generates a huge amount of data.
During the last few years, there was an explosion of the number of works 
at the interface of HPC and BigData, dealing with learning algorithms
(including the authors of this work~\cite{SC2015}).
Most of these studies target the process of determining more or less 
accurately the value of some specific key parameters, with the idea that
learning better estimates should improve the performances of the 
resource manager.

\bigskip
Our focus within this work is to learn how the management system behaves 
within the usual framework of selecting jobs of the submission queue 
under EASY-BF.
Contrarily to most existing studies, which consider only one execution 
trace, we extend the learning on a multiple traces basis.
We built a simulator which generates a huge number of partial executions 
coming from multiple users collected in various actual traces and HPC 
platforms.
The sampling is explained in detail later in the paper.
This was motivated because learning for a specific trace/platform leads 
to limited results, customized for one particular instance (then, what 
happens if we target another platform?).
The collected data are too small and moreover, this approach is usually 
inefficient to deal with extreme unpredictable events.
The sampling generation allow to smooth the data.

%On se place du point de vue de l'admin system...

@@

Our main contribution is to investigate a new way for learning, which 
consists in using of a simulator to generate a large data set extracted 
from various actual traces.
We propose two variants of the approach: the ideal situation which 
relies on the  precise knowledge of data, and the more realistic 
assumption of noisy data (which better reflect the uncertainties).
We also propose a new on-line approach based on multi-armed Bandit to 
learn very fast how to prioritize the submission and BackFilling queues.
\bigskip

Our main results are:

l'approche basee sur le simulateur marche bien. meme avec des donnes 
imprecises.

Le bandit a des resultats raisonnables, il ne necessite que peu de code 
supplementaire...

Dans les details :
on montre que une gestion aleatoire des queues est meilleure que FCFS, 
et on identifie des politiques qui marchent bien dans plusieurs traces, 
etc..

@@

This section presents and discuss the most significant results related 
to job scheduling and learning algorithms in this context.
Let us start by recalling some basics about scheduling heuristics in HPC 
platforms:

While parallel job scheduling is a old studied theoretical
problem~\cite{leung2004handbook} \texbf{PREFER DROZDOWSKI HERE + SURVEY 
FEITELSON-RUDOLPH-SWIEGELSHON}, the practical ramifications, varying
hypotheses, and inherent uncertainty of the problem applied on the HPC 
field have driven
practitioners to use \textit{simple} heuristics (and researchers to 
study their behavior).
The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.

While Conservative Backfilling offers many advantages~\cite{bfchar}, it 
has a
significant computational overhead, which mainly explains why most of 
the machines
of the top500 ranking~\cite{top500} rather use a variant of EASY-BF 
instead.
There is a large body of work seeking to improve/tune EASY. Indeed, 
while the
basic mechanism is used by some actual resource and job management 
softwares (most notably
SLURM~\cite{SLURMdocSCHED}), this is rarely done without fine tunings by 
system
administrators.

The original EASY mechanism refers to a First-Come-First-Serve basis.
Several works explore how to tune EASY by reordering waiting
and/or backfilling queues~\cite{Tsafrir_easypp_2005}, sometimes even in a
randomized manner~\cite{1592720}, as well as some
implementations~\cite{Jackson2001}. However, as successful as they may 
be, these works
do not address the dependency of scheduling metrics on the
workload~\cite{variability}. Indeed these studies most often report 
\textit{post-hoc} performance since
they compare algorithms after the workload is known.

The dynP scheduler~\cite{streit_selftuning_2002} proposes a systematic
method to tuning these queues, although it requires simulated scheduling 
runs
at decision time and therefore, it costs much more than the natural 
execution of EASY.

\paragraph
\textbf{Data-aware resource management.}

There was a recent focus on leveraging the huge amount of data available in
large scale computing systems in order to improve their behavior. Some works
use collaborative filtering to colocate tasks in clouds by estimating
application interference~\cite{7516031}.  Some others are closer to the 
application
level and use binary classification to distinguish benign memory faults from
application errors in order to execute recovery algorithms 
(see~\cite{fmodeling} for instance).

Several works use \textbf{this method} (LAQUELLE?) in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that 
better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.
However, these works do not address the duality between the cumulative and
maximal scheduling costs, as mentioned in~\cite{learningruntimes}.
While these previous works intend to estimate uncertain parameters, we 
consider here a more
pragmatic approach, which directly learn a good scheduling policy from a 
given policy space.

\paragraph
\textbf{Multi-Armed Bandits.}

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards. At every round, an action (represented by 
an arm of the Bandit) must be chosen
in a fixed set and the corresponding reward is observed. The goal of a MAB
algorithm is to maximize the total reward obtained in a successive 
number of rounds.
There exist a bunch of works that address this problem under a variety of
constraints, The two most popular settings are the original stochastic  case
and the adversarial case. See~\cite{thompson} for the original work on the
stochastic case and~\cite{Auer2002} for the UCB family of algorithms 
\textbf{EXPLAIN HERE THE ACRONYM}.
According to our knowledge, the work
of~\cite{Banos} is the earliest known work to us for the adversarial case
see~\cite{nonstoch} for the EXP family of algorithms. We refer to the
review~\cite{bubnow} for a comprehensive overview of the field. While these
algorithms bound the cumulative difference in loss to the best arm (the
regret), they have functional constraints such as the fact that the rewards
should be contained in a range. The heuristic called
\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to 
achieve good results in practice, and does not have
this requirement.

@@

%wait mwait lpf spf sqf lqf expfact mexpfact lrf srf laf saf
\begin{enumerate}
  \item FCFS: First-Come First-Serve, which is the widely used default policy~\cite{easy}.
  \item LCFS: Last-Come First-Serve.
  \item SPF: Smallest estimated Processing time $\widetilde{p_{j}}$ First ~\cite{bfchar}.
  \item LPF: Longest estimated Processing time First.
  \item LQF: Largest resource requirement $q_j$ First.
  \item SQF: Smallest resource requirement First.
  \item LEXP: Largest Expansion Factor First~\cite{bfchar}, where the expansion
    factor is defined as follows:
  \begin{equation} \frac{start_j - r_j + \widetilde{p_j}}{\widetilde{p_j}} \end{equation}
  where $start_j$ is the starting time of job $j$.
  \item SEXP: Smallest Expansion Factor First
  \item LRF: Largest Ratio $\frac{p_j}{q_j}$ First
  \item SRF: Smallest Ratio First
  \item LAF: Largest Area $ p_j \times q_j$ First
  \item SAF: Smallest Area First
\end{enumerate}

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[scale=0.6]{variability.pdf}
  \caption{Variability in the weekly average waiting time of EASY($P,20h$) using the KTH-SP2 trace (pre-processing described in subsection~\ref{sub:traces}) for
  a few select policies from~\ref{sub:th}.}
  \label{fig:variability}
\end{figure*}

@@

Readers interested in replicating the experiments (or a part thereof) are
invited to peruse to the 'artifacts description' appendix of this paper.  We
provide to persistent archives containing the simulator, the analysis code
and an automated workflow that replicates all experiments and figures
present in this paper.

@@

\begin{figure*}[h]
  \centering
  \includegraphics[scale=0.6]{CTC-SP2.pdf}
  \includegraphics[scale=0.6]{KTH-SP2.pdf}\\
  \includegraphics[scale=0.6]{CEA-Curi.pdf}
  \includegraphics[scale=0.6]{SDSC-BLU.pdf}\\
  \includegraphics[scale=0.6]{SDSC-SP2.pdf}
  \includegraphics[scale=0.6]{ANL-Intr.pdf}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.6]{UniLu-Ga.pdf}

@@

  \caption{Evolution of the average cumulative waiting time improvement
    compared to EASY-FCFS of the policies $P_i$ for $i = 1 \ldots 10$ on the
    UniLu-Gaia trace. The
    average is obtained by resampling the original trace 100 times. The dashed
    lines represent the 10th and 90th percentiles of the values across this
  resampling.}

@@

\clearpage

\section*{Artifacts Description Appendix}

This appendix describes the computational artifacts associated with this work. We
take a lightweight approach to replicability by packaging a declarative
workflow that generates all the figures from this paper and making it available
publicly. Moreover, questions regarding replication can be freely directed at
the authors via e-mail.

@@

There are various approaches for making computational experiments replicable,
among which distributing complete operating system images, containers, or
packaging software. As our experiments are CPU bound, we decide to opt for the
software packaging approach using the Nix~\cite{nix} packaging system. This
allows for replicating the experiments on clusters where virtualization is not
available or kernels are old and decreases runtime. All the dependencies of our
experiments (including our own code, dependencies for data processing and
visualization and workflow engine) are automatically managed by Nix up to the
actual execution of the code.

Running the experiments can be done on any platform equipped with the nix
package manager by running:

\begin{lstlisting}
nix-build (fetchTarball ...)
\end{lstlisting}

@@

 The experiments in this paper can run via in less than 24 hours on a moderately parallel
 host. More precisely, we had access to a Dell PowerEdge R730 equipped with
 a total of 56 threads @2.4GHz each and 757G of RAM. For this reason, the
 experimental campaign workflow was designed to be executed on a single host.

@@

 All experiments are tied together using Zymake~\cite{zymake}, a minimalistic
 workflow system designed for computational experiments\footnote{The zymake system
 is also packaged by our Nix expressions.}. This system is
 analogous to a traditional build system with added workflow capabilities. The
 entire workflow that generates this article from the input data is contained
 in a single \lstinline[basicstyle=\ttfamily\color{blue}]|zymakefile| to be
 found at the root of the main archive. This workflow is composed of the
 following steps:

 \begin{itemize}
   \item[Preprocessing]: Data extraction from archives, filtering, resampling.
     This is principally using shell scripts for data manipulation (see file
     \lstinline[basicstyle=\ttfamily\color{blue}]|misc/strong\_filter| for the
     filtering steps used) and ocaml code that implements the resampling method
     described in~\ref{ssub:resampling}.

@@

   \item[Analysis]: This step runs R code that generates the figures
     presented in this paper.
 \end{itemize}
\end{document}
